---
title: "ESM 206 Lab 10"
author: "Ashley S"
date: "December 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setRepositories(ind=1:2)
devtools::install_github("GuangchuangYu/ggimage")
devtools::install_github('thomasp85/gganimate')

```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images


```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}
#read in data
env_var <- read_csv("world_env_vars.csv") %>% 
  na.omit # going to delete any row that has na it will delete it, not recommended but doing it

#caclculate Pearson's e for correlations 
cor_df <- cor(env_var[2:28]) #if want all columns, just data frame, but we will say what columns we want here but putting in brackets. now we see same columns as row which is why diagnol is 1, bc they are perfectly aligned. range of values from really negative to really positive values. but this is a tedious way to view, easier to look at visually so cor plot function. 

#make a correlation matrix

corrplot(cor_df,
        type = "upper",
        method = "ellipse",
        tl.col = "black",
        tl.cex = 0.5) # do not put original data frame, you put your matrix data, elipse direction is directionallity of..., change . a really dark red band is negative correlation, see 13 minutes for explanation. 
# big pic view of how variable related to eachother. if you have categorical factors with different levels not super interesting since they can only be a few things. usually for continuous predictor variables. 


```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}
# interpretting coefficients in logodds model is complicated, so how run model and how find odds and then have r convert to probabilities that is more useful for an audience. 
#what effect do age and sex have on survival of donner party.

donnertable <- read_csv("DonnerTable.csv") # survivor is 1, death is zero and then the rest. when outcome variable already coded as 0 and 1, then there is no question about outcome a or b. in this case saying probability outcome of 0 is outcome of not survvial and prob of 1 is survival. YOu can specify with casewhen what is 1 and what is 0 so there is no confusion of what you are r think are reference. to avoid confusion of what r is choosing is you specify what a 0 outcome is and 1 outcome is. in this case it is specified for us bc allison told us. 





```

b. Binomial logistic regression

```{r blr}

donner_blr <- glm(Outcome ~ Sex + Age, family = "binomial", data = donnertable) #must tell them type of log linear model using/creating, so here binomial, 

summary(donner_blr)
#see both coeeficients for sex male and age are both negative which makes me think even though log odds, but based on DP outcomes, male survivorship is lower and increase in age lower survivorship. so what does this actually mean in terms of odds of survivorship?

```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 
# use predict function but feed it new data to make predictions for and must have exact same variable names for the data model is looking for when it is created. 

f_5 <- data.frame(Sex = "Female", Age = 5) #so exact name and case her, otherwise would not be compatible bc variables would be mismatched. 


# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":
# now new DF, so now feed into model using predict function
f_5_logodds <- predict(donner_blr, newdata = f_5, type = "link")
f_5_logodds # 1.443775, if want to convert, exponentiate it. 

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

f_5_odds <- exp(f_5_logodds)
f_5_odds # 4.23, so 4:1, so probability of survival is 80%, so prob of it happening divided by prob of it not happening. 


# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS
# 

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f_5_prob <- predict(donner_blr, newdata = f_5, type = "response")
f_5_prob # so we get .809, so this is exactly what we found, way easier to digest than reporting logodds, in a peer reviewed you would report logit, but not for general audience. 

```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donner_blr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob

# press play button in corner to run the whole thing. see it is 42% . this makes sense bc women and children stay put for help while men went out to look for help. 

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

# this code chunk is to create a DF of new data to feed into that model to make predictions that has a sequence of ages for sex to make predictions for. first makes a vector called sequence age which repeats ages for 0 -100 twice. notice 0 - 100 is 101 values. so next vector repeat string female 101 times and male 101 times and then bind together into a single vector. so ages 0-100 twice and a series of strings where 101 male repeated 101 female repeated then bind together so columns side by side. the problem is model will not recognize seq_age and mf, it only recognizes age and sex, so you have to update column names. there are many ways to do this, some more eloquent than this. 

#look at donner new data frame, 


```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.
#now feed in data frame with females and males

predicted_probs <- predict(donner_blr, newdata = donner_newdata, type = "response", se.fit = TRUE) # se calculates se for prob of survival, use response bc looking at survival. only makes sense to view once bind together.  

# Coerce outcome into data frame. 

graph_data <- data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)
colnames(graph_data) <- c("Age", "Sex", "Probability", "SE") #could also use rename 
#please make DF that uses the columns then bind the predicted probability which is stored as fit in predicted probs, then another column of SE
#make a line graph that shows prediction of probability of our model then ribbon to show plus minus SE
```

g. Graph results.

```{r donner_graph}

ggplot(graph_data, aes(x = Age, y = Probability)) +
  geom_line(aes(color = Sex)) +
  geom_ribbon(aes(ymin= Probability - SE, ymax= Probability + SE, fill = Sex), alpha = 0.4)
#R has no idea that there is a difference between male and female, so need to add factor to create different series, can do aesthetically that color denotes sex which tells r there is different series you want to exist. with any model some measure of uncertainty, so add ribbons around predictions that will show the SE. when you create a ribbon plot is manually saying what upper and lower bounds should be of that ribbon width. so at any point the max value of that ribbon is probability plus 1 SE and the min value is probability minus SE. 
#now graph shows predicted prob of males and females and ribbon width shows SE
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}
# read in 3 files into the same data frame. if we have multiple files with sim formatting and same column names, especially if something similar about file prefix. so consistency in format of actual data and file naming system. 

si_full <- list.files(pattern = "solar_irradiation_*") %>% 
  map_df(~read_csv(.)) %>% 
  clean_names()
# first make a list of complete file names with anything that starts with solar radiation . asterisk at the end says anything after works but prefix will be all this in a list. map df is put all in a df and ~ is a function to say put these all together, period is outcome of previous line. it is problematic if there is not an indicator that specifies the levels of that different data entered. since each had site alaska, site sb, etc that identifier is what we needed. but notice, these column names are terrible, so could change names in escel, could use rename in R, but janitor package has cleannames will auto change everything in the snake case for you including case. Snake case is only lowercase letters and any space in punctation is replaced with an underscore. 

# if you look at class, R understands as a charachter not as a date and time. so use lubridate. 


```

b. Wrangle the data
```{r wrangle_solar}

solar_tidy <- si_full %>% 
  rename(sol_rad = etr_wh_m_2,
         date = yyyy_mm_dd,
         time = hh_mm_lst) %>% 
  filter(time != "NA") %>% 
  mutate(site = fct_relevel(site, "Hawaii", "Santa Barbara", "Alaska"))
# remember != means remove, so explicity set factor levels by this relevel since auto by alphabetically
```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

solar_tidy$date <- mdy(solar_tidy$date)
solar_tidy$time <- hms(solar_tidy$time)
#so now reformat in the class of month datae year of what currently exists in date. it makes logical assumptions of what could be a year month date etc, and assigns the class, if there are conflicts it will let you know then manually specify, for example 2018 in a column it will know a year. if now run class(colar_tidy$date) in console it was now show it is now a date

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}
solar_gg <- ggplot(solar_tidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) +
  scale_fill_gradientn(colors = c("royalblue2", "mediumorchid1", "yellow", "orange"))+
  scale_y_time()+
  facet_wrap(~site, ncol = 1)

solar_gg

#color of tile will be based on that vaalue. so make ggplot. you have to tell the scale of axis is time, so add scale. but still need site differences across sites, so add facet wrap. if want separated for 1 column, do ncol = 1, so columns are own separate row. scale fill gradient n allows you to add as many colors as you want in your scheme.
# heat map good for value of variable where 2 other variables inform about that variable. 

```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}

aq_df <- read_csv("aq_wb.csv")
# data is in wide format, not tidy bc would have one column of year and different years for different countries in one column. 

aq_tidy <- aq_df %>% 
  filter(country == "Brazil"|
           country == "Chile"|
           country == "Ecuador"|
           country == "United States") %>% 
  gather(year, aq_prod, `1960` : `2016`) %>% 
  filter(year >= 1990) %>% 
  mutate(aq_mil = aq_prod/1000000) %>% 
  select(country, year, aq_mil)
  

#if in wide format but want in long format, use function gather. so list the 2 columns creating, then info that will be stored in those columns. it will take all of column headers and put into first column with correct number of repitions for ..., then all values for those columns and attach in a separate column for aq prodution . so give it all columns you are compressing into the two columns, since numbers use backticks around it, tells r these are column names and not a sequence of values. convert amount column to millions of metrics tons, so mutate. 
```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish <- "fish.png"

```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aq_plot <- ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country ))+
  geom_line(aes(color = country))+
  geom_image(aes(image = fish)) +
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x = 1), size = 5)+
  transition_reveal(country, as.numeric(year))



aq_plot

# can export as a gif by animate(aq_plot, nframe = 24, renderer = gifski_renderer("aq_animate.gif"))

# when gathered stil thinks it is characters so coerce it here, default for text lable is right over center of point, so we want to be above and shifted to the right, so position. 

```

##END LAB